---
title: "Data Visualization Workshop"
author: "Sports Analytics Group"
date: "2024-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This workshop will preview some basic data visualization techniques in R using the `ggplot2` library and assumes some prior knowledge and experience with the `tidyverse` library. For more extensive notes on data manipulation, see [here](https://github.com/jeremydumalig/Sports-Analytics-Group).

## Brief Review of Data Manipulation using `tidyverse`

Load package and data (this dataset and others can be found [here](https://github.com/jeremydumalig/Sports-Analytics-Group)).

```{r, message=F}
library(tidyverse)
```

```{r}
nba = read_csv("nba23.csv")

nba %>%
  mutate(MIN = MP / G,
         `AST/TO` = AST / TOV) %>%
  filter(Tm == "DEN",
         MIN >= 10) %>%
  arrange(desc(`AST/TO`)) %>%
  select(Player, Pos, AST, TOV, `AST/TO`) %>%
  head(5)

nba %>%
  group_by(Tm) %>%
  summarize(`3PM` = sum(`3P`),
            `3PA` = sum(`3PA`)) %>%
  ungroup() %>%
  mutate(`3P%` = 100 * `3PM` / `3PA`) %>%
  arrange(`3P%`) %>%
  head(5)
```

## Data Visualization using `ggplot2`

Documentation for plotting in `ggplot2` found [here](https://rstudio.github.io/cheatsheets/html/data-visualization.html).

```{r, message=F}
library(ggplot2)
```

```{r, warning=F, fig.width=10}
df = 
  nba %>%
  mutate(`FG%` = 100 * FG / FGA)

ggplot(df) +
  geom_point(aes(x=PTS, 
                 y=`FG%`,
                 color=Pos),
             size=3,
             alpha=0.75) +
  theme_bw()

nba %>%
  ggplot() +
  geom_histogram(aes(x=`3P`),
                 color='darkblue',
                 bins=30) +
  labs(x="3-Point Makes (3PM)",
       y="Count",
       title="Distribution of Player 3PM in the 2022-23 NBA Season") +
  theme_bw()

nba %>%
  filter(Tm %in% c("GSW", "LAL", "LAC", "SAC", "PHO")) %>%
  group_by(Tm) %>%
  summarize(`3PM` = sum(`3P`),
            `3PA` = sum(`3PA`)) %>%
  ungroup() %>%
  mutate(`3P%` = 100 * `3PM` / `3PA`) %>%
  ggplot() +
  geom_bar(aes(x=Tm, 
               y=`3P%`,
               fill=Tm),
           stat='identity') +
  lims(y = c(0, 50)) +
  theme_bw()
```





